---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence? 
A1. Contributions made 10 and 25 dollars were the most frequent (5674), contributions of 500 dollars or greater were the least frequent (57).

```{r}
wv_winred_categories <- wv_winred_contribs %>%
  select(last_name, first_name, city, date, amount, memo_text) %>%
  mutate (
    amount_category = case_when(
      amount < 10 ~ "Less than $10",
      amount >= 10 & amount < 26 ~ "Between $10 and $25",
      amount >= 26 & amount < 51 ~ "Between $26 and $50",
      amount >= 51 & amount < 101 ~ "Between $51 and $100",
      amount >= 101 & amount < 500 ~ "Between $100 and $500",
      amount >= 500 ~ "$500 and up"
      
    ) 
  ) %>%
   group_by(amount_category) %>% 
    summarise(total_contribs = n()) %>% 
    arrange(desc(total_contribs))
  
  
```
Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  
A2. Dr. John Almond, Franco Colonna and Betty Bender were the top 3 donors of contributions less than $10 in WV. 

```{r}
standardized_wv_contribs <- wv_winred_contribs %>% 
  mutate(
    upper_last_name = str_to_upper(last_name),
    upper_first_name = str_to_upper(first_name)
  ) 
```
```{r}
standardized_wv_contribs <- standardized_wv_contribs %>% 
  select(upper_last_name, upper_first_name, amount, memo_text) %>% 
  mutate(
    amount_category = case_when(
      amount < 10 ~ "Less than $10",
      amount >= 10 ~ "$10 and up"
    )
  )
```
```{r}
standardized_wv_contribs <- standardized_wv_contribs %>% 
  filter(amount_category == "Less than $10") %>% 
  unite('full_name', upper_first_name, upper_last_name, sep=' ')
```
```{r}
top_small_donors <- standardized_wv_contribs %>% 
  count(full_name) %>% 
  arrange(desc(n))
#for some reason the code blocks would create an error when i combined them into one giant code block but it worked separately
```
Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.
A3. Why did so many people make constant small donations? Were they all to different people? One person? Were the donations they made really small, like less than $1? Or were they several dollar donations?

Q4. Which month saw the most contributions?
A4. 

```{r}
intall.packages('lubridate')
```
```{r}
#whenever i would type a function for lubridate, an error message would pop up, saying that the lazy-load database fille was corrupt, also i wasn't sure how to transfer the lubridate part we learned in the pre_lab to this question
```
Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.
A5. There were six donations of $2900. Half of them went to Carol Miller who is an incumbent congress representative for WV. One contribution went to David McKinley who is also an incumbent representative for WV. One donation is for Steven Palazzo, an incumbent representative for Mississippi. The last donation was for Ryan Zinke, who has political experience but is running for U.S. congress in Montana for the first time.

```{r}
total_top_donations <- wv_winred_contribs %>% 
  select(date, amount, memo_text) %>% 
  filter(amount == 2900) 
  
```
