---
title: "lab_01"
author: "Derek Willis"
date: "7/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
wv_winred_contribs <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** Rows: 13,382 Columns: 24

```{r}
glimpse(wv_winred_contribs)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?
**A2** Martinsburg with 781

```{r}
```count_contribs_city = wv_winred_contribs %>% group_by(city) %>% summarise(count_contribs_city = n())
  glimpse(count_contribs_city)

**Q3** Which West Virginia city accounted for the fewest contributions? How many?
**A3** 40 different cities each had 1 contribution, some city names include Auburn, Belmont, Camden on Gauley, New Milton, Smoot and Unger.

```{r}
`count_contribs_city = wv_winred_contribs %>% group_by(city) %>% summarise(count_contribs_city = n())
  glimpse(count_contribs_city)

**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** Martinsburg is the biggest city in the eastern panhandle and 8th biggest in all of W.V. with over 18,000 residents. The other towns are all much smaller, rural towns. With more residents and being in the eastern panhandle that's closer to D.C. and Baltimore, there are more people with likely higher paying jobs, creating more contributions.

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?
**A5** The Save America Joint Fundraising Committee recieved the most money. The total number of contributions was 2448, the total money was $66972.89 and the mean contribution was $27.358207
 
```{r}wv_winred_contribs %>%
+     group_by(memo_text) %>%
+     summarise(
+         count_contribs = n(),
+         total_amount = sum(amount),
+         mean_amount = mean(amount)
+     ) %>%
+     arrange(desc(total_amount))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** The Save America Joint Fundraising Committee is affiliated with former President Donald Trump, and people tend to care most about presidential elections and campaigning the most.

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by city):
* count of contributions
* total amount of all contributions
* the average contribution amount
* the median contribution amount
* the smallest contribution amount
* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7**. The average contributor to the republican party in W.V. elections donated $26.66, a good comparison would be how much the average W.V. contributor paid to the democratic party. The same could be done with the total amount donated.

Another comparison story could be how much is being donated to the republican party in W.V. compared to other states. This could give insight to the socioeconomic status of people in different states and how willing they are to donate to their party of interest.

Another interesting story could be why there is such a wide discrepancy in the smallest contribution in $0.01 to the largest contribution in 5800. 



```{r}
wv_summary = wv_winred_contribs %>%
+     summarise(
+         count_contribs = n(),
+         total_amount = sum(amount),
+         mean_amount = mean(amount),
+         median_amount = median(amount),
+         min_amount = min(amount),
+         max_amount = max(amount)
+     )
```
